---
title: "Fundamentals of Computing and Data Display"
author: "Christoph Kern"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    df_print: kable
    highlight: textmate
    theme: simplex
    toc: yes
subtitle: R Markdown and LaTeX
bibliography: rmarkdown.bib
---

```{r, include = FALSE}
library(knitr)
library(RSocrata)
library(tidyverse)
```

## Data
***
This notebook exemplifies some features of R Markdown [@Xie2018]. We use the [Detroit crime data](https://dev.socrata.com/foundry/data.detroitmi.gov/9i6z-cm98) as our example data set. Note that this code chunk utilize caching, i.e. the actual API query will be executed only once.

```{r, cache = TRUE}
url <- "https://data.detroitmi.gov/resource/9i6z-cm98.json?$where="
d18 <- read.socrata(paste0(url,"incident_timestamp between '2018-01-01' and '2018-07-01'"))
```

Data tables can be formatted in different styles. We use the `knitr::kable` function as an option in the YAML header of this document.

```{r}
d18$hour_of_day <- as.numeric(d18$hour_of_day)
d18[1:5, 1:5]
```

## Example code
***
Inline code can be used to directly display code results in the text. For example, the number of crime incidents in our data is `r nrow(d18)`.

There are a number of chunk options to control the layout of figures. In addition, we can hide the code itself with `echo = FALSE`.

```{r, echo = FALSE, message = FALSE, fig.align = "center", fig.cap = "Crime in Detroit", out.width = "55%"}
d18 %>%
  separate(incident_timestamp, c("date", "time"), sep = " ") %>%
  filter(as.Date(date) < "2018-07-01") %>%
  group_by(date) %>%
  summarise(total = n()) %>%
  ggplot() +
  geom_line(aes(x = as.Date(date), y = total)) +
  geom_smooth(aes(x = as.Date(date), y = total)) +
  ylim(0, 400) +
  labs(x = "Date", 
       y = "Total") +
  theme_light()
```

## References
